Supplementary material

### 2021-07-12 submission/manuscript.rmd
cross-validate and replicate 100 times for gusto evaluation

```{r, echo=FALSE, warning=FALSE, message=FALSE}
gusto <- gusto %>%
  tibble() %>%
  rename(
    "outcome" = "day30",
    "treatment" = "tpa"
  )

prediction <- lrm(
  outcome ~ treatment + age + Killip + pmin(sysbp, 120) + lsp(pulse, 50) + pmi + miloc,
  data = gusto,
  maxit = 99
)

riskLinearPredictor <- predict(
  prediction,
  newdata = gusto %>% mutate(treatment = 0) %>% data.frame()
)

gusto <- gusto %>%
  mutate(
    riskLinearPredictor = riskLinearPredictor
  )

constantModel <- fitModelBasedHte(
  data     = gusto, 
  settings = createModelBasedSettings(type = "treatment")
)

tmp <- gusto %>%
  mutate(
    predictedBenefit = predictBenefitModelBasedHte(
      p             = plogis(.$riskLinearPredictor), 
      modelBasedFit = constantModel
    )
  )

constantC   <- calculateCForBenefit(tmp)
constantIci <- calculateCalibrationForBenefit(tmp)

stratified <- fitStratifiedHte(
  data     = gusto,
  settings = createStratifiedSettings()
)

linearModel <- fitModelBasedHte(
  data     = gusto,
  settings = createModelBasedSettings()
)

tmp <- gusto %>%
  mutate(
    predictedBenefit = predictBenefitModelBasedHte(
      p             = plogis(.$riskLinearPredictor), 
      modelBasedFit = linearModel
    )
  )

linearC   <- calculateCForBenefit(tmp)
linearIci <- calculateCalibrationForBenefit(tmp)

rcs3Model <- fitRcsHte(
  data     = gusto,
  settings = createRcsSettings()
)

tmp <- gusto %>%
  mutate(
    predictedBenefit = predictSmoothBenefit(
      p          = plogis(riskLinearPredictor), 
      smoothFit  = rcs3Model
    )
  )

rcs3C   <- calculateCForBenefit(tmp)
rcs3Ici <- calculateCalibrationForBenefit(tmp)

rcs4Model <- fitRcsHte(
  data     = gusto,
  settings = createRcsSettings(nKnots = 4)
)

tmp <- gusto %>%
  mutate(
    predictedBenefit = predictSmoothBenefit(
      p          = plogis(riskLinearPredictor), 
      smoothFit  = rcs4Model
    )
  )

rcs4C   <- calculateCForBenefit(tmp)
rcs4Ici <- calculateCalibrationForBenefit(tmp)

rcs5Model <- fitRcsHte(
  data     = gusto,
  settings = createRcsSettings(nKnots = 5)
)

tmp <- gusto %>%
  mutate(
    predictedBenefit = predictSmoothBenefit(
      p          = plogis(.$riskLinearPredictor), 
      smoothFit  = rcs5Model
    )
  )

rcs5C   <- calculateCForBenefit(tmp)
rcs5Ici <- calculateCalibrationForBenefit(tmp)

gustoPerformance <- tibble(
  discrimination = c(
    constantC, linearC, rcs3C,
    rcs4C, rcs5C
  ),
  calibration = c(
    constantIci$ici, linearIci$ici, rcs3Ici$ici,
    rcs4Ici$ici, rcs5Ici$ici
  ),
  method = c(
    "constant treatment effect",
    "linear interaction",
    "RCS smoothing with 3 knots",
    "RCS smoothing with 4 knots",
    "RCS smoothing with 5 knots"
  )
)

cBenefit <- gustoPerformance %>% arrange(discrimination)

```

We used the risk linear predictor to fit the proposed methods under study for
individualizing absolute benefit predictions. All methods predicted increasing
benefits for patients with higher baseline risk predictions, but the fitted
patterns were clearly different. The adaptive approach selected the model with
RCS smoothing with 4 knots. However, for very low baseline risk this model
predicted decreasing benefit with increasing risk maybe somewhat too flexible.
The more robust models, the linear interaction model or the model with RCS
smoothing (3 knots), gave very similar benefit predictions, followed the
evolution of the stratified estimates very closely and may therefore be
preferable for use in clinical practice.

>>>>

```{r, echo=FALSE, warning=FALSE, message=FALSE}
gustoPerformance <- readr::read_csv(here::here("data/processed/gustoPerformanceMetrics.csv"))

discriminationPerformance <- gustoPerformance %>%
  select("discrimination")
calibrationPerformance <- gustoPerformance %>%
  select("calibration")
```

We used the risk linear predictor to fit the proposed methods under study for
individualizing absolute benefit predictions. All methods predicted increasing
benefits for patients with higher baseline risk predictions, but the fitted
patterns were clearly different. The adaptive approach selected the model with
RCS smoothing with 4 knots. However, for very low baseline risk this model
predicted decreasing benefit with increasing risk may be somewhat too flexible.
The more robust models, the linear interaction model or the model with RCS
smoothing (3 knots), gave very similar benefit predictions, followed the
evolution of the stratified estimates very closely and may therefore be
preferable for use in clinical practice. The linear interaction model
had somewhat lower AIC compared to the model with RCS smoothing (3 knots), slightly better
cross-validated discrimination (c-for-benefit
`r round(discriminationPerformance[2, ], 3)` vs
`r round(discriminationPerformance[3, ], 3)`) and quite similar cross-validated
calibration (ICI-for benefit `r round(calibrationPerformance[2, ], 4)` vs
`r round(calibrationPerformance[3, ], 4)`).

### 2021-07-12 submission/manuscript.tex
cross-validate and replicate 100 times for gusto evaluation

We used the risk linear predictor to fit the proposed methods under
study for individualizing absolute benefit predictions. All methods
predicted increasing benefits for patients with higher baseline risk
predictions, but the fitted patterns were clearly different. The
adaptive approach selected the model with RCS smoothing with 4 knots.
However, for very low baseline risk this model predicted decreasing
benefit with increasing risk may be somewhat too flexible. The more
robust models, the linear interaction model or the model with RCS
smoothing (3 knots), gave very similar benefit predictions, followed the
evolution of the stratified estimates very closely and may therefore be
preferable for use in clinical practice. The model with RCS smoothing (3
knots) had somewhat lower AIC compared to the linear interaction model,
slightly better cross-validated discrimination (c-for-benefit 0.537 vs
0.527) and quite similar cross-validated calibration (ICI-for benefit
0.0114 vs 0.0116).

>>>>

We used the risk linear predictor to fit the proposed methods under
study for individualizing absolute benefit predictions. All methods
predicted increasing benefits for patients with higher baseline risk
predictions, but the fitted patterns were clearly different. The
adaptive approach selected the model with RCS smoothing with 4 knots.
However, for very low baseline risk this model predicted decreasing
benefit with increasing risk may be somewhat too flexible. The more
robust models, the linear interaction model or the model with RCS
smoothing (3 knots), gave very similar benefit predictions, followed the
evolution of the stratified estimates very closely and may therefore be
preferable for use in clinical practice. The linear interaction model
had somewhat lower AIC compared to the model with RCS smoothing (3
knots), slightly better cross-validated discrimination (c-for-benefit
0.526 vs 0.525) and quite similar cross-validated calibration (ICI-for
benefit 0.0115 vs 0.0117).

### 2021-07-12 submission/supplement.rmd
cross-validate and replicate 100 times for gusto evaluation

---
title: "Supplementary material"
output:
    bookdown::pdf_document2: default
geometry: margin=1.0in
toc: false
font-size: 12pt
header-includes:
  - \usepackage{amssymb}
  - \usepackage{bm}
  - \usepackage{booktabs}
  - \newcommand\given[1][]{\:#1\vert\:}
  - \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}

editor_options: 
  chunk_output_type: console
bibliography: references.bib
csl: jce.csl
---

@@

 knitr::kable(
    metrics,
    format      = "latex",
    align       = "l",
    booktabs    = TRUE,
    digits      = 4,
    format.args = list(
      big.mar      = ",",
      decimal.mark = "."
    ),
    col.names = c(
      "Model",
      "C",
      "ICI"
    ),
    caption    = "Performance metrics of the different methods used to predict benefit of tPA on 30-day mortality using data from the GUSTO-I trial."
  ) 
```

>>>>

---
title: "Supplementary material"
output:
    bookdown::pdf_document2: default
geometry: margin=1.0in
toc: false
font-size: 12pt
header-includes:
  - \usepackage{amssymb}
  - \usepackage{bm}
  - \usepackage{booktabs}
  - \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
  - \newcommand\given[1][]{\:#1\vert\:}
editor_options: 
  chunk_output_type: console
bibliography: references.bib
csl: jce.csl
---

@@

 knitr::kable(
    metrics,
    format      = "latex",
    align       = "l",
    booktabs    = TRUE,
    digits      = 4,
    format.args = list(
      big.mar      = ",",
      decimal.mark = "."
    ),
    col.names = c(
      "Model",
      "C",
      "ICI",
      "AIC"
    ),
    caption    = "Performance metrics of the different methods used to predict benefit of tPA on 30-day mortality using data from the GUSTO-I trial."
  ) 
```
